---
title: "Training Day Jan 17"
author: "Matthew Druckenmiller"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, echo = F, message = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
```


Create a dataframe of the fish escapement data from <https://knb.ecoinformatics.org/view/doi:10.5063/F1S46Q6M>

```{r, echo = F}
esc <- read.csv(file="https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", header=TRUE, stringsAsFactors = F)
#head(esc)
```

Calculate annual, regional total escapement

* `separate` the date into year - month - day
* `group-by` `summarize` to calculate the sum over unique region-year-species combinations
* `filter` for salmon species

```{r, echo = F}
annual_esc <- esc %>%
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.numeric(year)) %>%
  group_by(SASAP.Region, Species, year) %>%
  summarise(escapement = sum(DailyCount)) %>%
  filter(Species %in%  c("Chinook", "Sockeye", "Pink", "Chum", "Coho"))
#head(annual_esc)
```

```{r, echo = F}
datatable(annual_esc)
```

# Static Plots

You can also create a common theme to use for all:
```{r}
my_theme <- theme_bw() + theme(plot.title = element_text(hjust=0.5))
```


```{r, echo=F}
ggplot(annual_esc, aes(x= Species, y = escapement, fill=SASAP.Region)) +
  geom_col() + # creates the columns +
  my_theme
```

Timeseries of Kodiak escapement by species

```{r}
ggplot(data=filter(annual_esc, SASAP.Region == "Kodiak"),
       aes(x = year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  ylab("Escapement (count)")+
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
  #theme_bw() +
  #theme_minimal()
  #theme_dark()
  #theme(plot.title = element_text(hjust=0.5))
```

Time series

```{r, fig.height=10, fig.width=4}
p <- ggplot(annual_esc, aes(x = year, y = escapement, color = Species))+
  geom_line()+
  geom_point()+
  facet_wrap(~SASAP.Region, scales="free", ncol=1)+
  my_theme
p
ggsave("region_escapement.png", p, width=6, height=12)
```

