---
title: "ADC-training-tidying"
author: "Matthew Druckenmiller"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F}
# chunk for library calls:
library(dplyr)
library(tidyr)
```

# Data Cleaning

Read in data file

```{r}
catch <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1",
                  method="libcurl"),
                  stringsAsFactors = FALSE)
# the libcurl method is related to my operating system and how it reads a url
head(catch)
```

We are going to use the pipe operator %>% in order to:  

* remove the sum and notes columns  
* convert from wide to long format  

```{r}
catch_long <- catch %>%
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) %>%
  # or select(-All, -notesRegCode)
  gather(key="species", value="catch", -Region, -Year)
  # gather transforms the data to a long format (spread would do the opposide)
head(catch_long)
```

Next, we use the pipe to:

* rename the catch column as "catch-thousands"  
* remove non-numeric characters  
* add a catch column with full values (not factors of a 1000)  

```{r}
catch_cleaned <- catch_long %>%
  rename(catch_thousands = catch) %>%
  mutate(catch_thousands=ifelse(catch_thousands=="I",1,catch_thousands)) %>%
  mutate(catch_thousands=as.numeric(catch_thousands)) %>%
  mutate(catch=catch_thousands*1000)
head(catch_cleaned)
```

```{r, eval=F, echo=F}
#We used to locate any non-numeric values in the dataframe's catch_thousands
test_catch <- as.integer(catch_cleaned$catch_thousands)
i <- which(is.na(test_catch)==T)

catch_cleaned[i,]
```

# Split, apply, combine

Calcuate total catch by region:

```{r}
catch_total <- catch_cleaned %>%
  group_by(Region, Year) %>%
  summarize(catch_region = sum(catch),
            n_obs=n())
catch_total
```

Calcuate the average fish by species:

```{r}
catch_species <- catch_cleaned %>%
  group_by(species, Year) %>%
  summarize(catch_mean = mean(catch), catch_sd = sd(catch), n_obs=n())
catch_species
```

Filter for chinook:

```{r}
catch_chinook <- catch_cleaned %>%
  filter(species=="Chinook" | species=="Chum") %>%
  arrange(-Year)
head(catch_chinook)
```

# Joins

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                        method = "libcurl"),
                        stringsAsFactors = FALSE)
head(region_defs)
```

```{r}
region_clean <- region_defs %>%
  select(code, mgmtArea)
head(region_clean)
```

We will join below by specifying the columns to join by:

```{r}
catch_joined <- left_join(catch_cleaned, region_clean, by=c("Region"="code"))
head(catch_joined)
```


